---
title: "fittingModels"
author: "Jacob Mitchell"
date: "4/15/2022"
output: html_document
---
## Read in the Data

```{r setup}
rm(list = ls())
source("/Users/jacobmitchell/Documents/STAT6440/functions/myfunctions.R")
hospital <- read.csv("/Users/jacobmitchell/Documents/STAT6440/csvs/dataset_mice_output.csv")
hospital$hospital_death <- as.factor(hospital$hospital_death)
levels(hospital$hospital_death) <- c("0", "1")
hospital = subset(hospital, select = -c(apache_4a_hospital_death_prob, apache_4a_icu_death_prob))
```
## Data Munging
Changing non numerical factors to be factors, and convert ethnicity, admit source, etc to numerical factors. 
apache_4a_hospital_death_prob, apache_4a_icu_death_prob were dropped in the prediction process since we don't have the model that created these probablities, and want to see what other factors play a role in prediction of death.
```{r dataCleaning}
vals <- unique(hospital$ethnicity)
hospital$ethnicity <- as.factor(hospital$ethnicity)
hospital$ethnicity <- ifelse(hospital$ethnicity == "Caucasian", 1, hospital$ethnicity)
hospital$gender <- as.factor(hospital$gender)
hospital$gender <- ifelse(hospital$ethnicity == "M", 0, hospital$gender)
hospital$gender <- ifelse(hospital$ethnicity == "F", 1, hospital$gender)
hospital$icu_admit_source <- as.factor(hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Accident & Emergency", 0, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Operating Room / Recovery", 1, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Floor", 2, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Other Hospital", 3, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Other ICU", 4, hospital$icu_admit_source)
hospital$icu_stay_type <- as.factor(hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "admit", 0, hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "readmit", 1, hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "transfer", 2, hospital$icu_stay_type)
hospital$icu_type <- as.factor(hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "Med-Surg ICU", 0, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "MICU", 1, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "CTICU", 2, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "CCU-CTICU", 3, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "Neuro ICU", 4, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "SICU", 5, hospital$icu_type)
hospital$apache_3j_bodysystem <- as.factor(hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Cardiovascular", 0, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Metabolic", 1, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Nuerological", 2, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Gastrointestinal", 3, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Metabolic", 4, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Musculoskeletal/Skin", 5, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Respiratory", 6, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Trauma", 7, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Genitourinary", 8, hospital$apache_3j_bodysystem)
hospital$apache_2_bodysystem <- as.factor(hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Cardiovascular", 0, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Metabolic", 1, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Nuerological", 2, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Gastrointestinal", 3, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Metabolic", 4, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Musculoskeletal/Skin", 5, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Respiratory", 6, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Trauma", 7, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Genitourinary", 8, hospital$apache_2_bodysystem)
```
## Data Partition on MICE imputed dataset

```{r dataPartition}
RNGkind(sample.kind = "Rounding")
set.seed(0) ## set seed so that you get same partition each time
p2 <- partition.2(hospital, 0.7) ## creating 70:30 partition
training.data <- p2$data.train
test.data <- p2$data.test
```

## Fit Classification Tree on MICE imputed Dataset

```{r pressure}
library(rpart)
library(rpart.plot)
library(caret)
set.seed(0)
train_control <- trainControl(method="cv", number=10)
cv.ct <- train(hospital_death ~ . , data = training.data, method = "rpart",trControl = train_control, tuneLength = 10,metric="Kappa")
```

```{r plottingTree}
print(cv.ct)
plot(cv.ct)
```

```{r testing}
cv.ct$finalModel
prp(cv.ct$finalModel, type = 1, extra = 1, under = TRUE, split.font = 2, varlen = -10)
```

```{r resultsOfTree}

pred.test.prune = predict(cv.ct$finalModel, test.data, type = 'class')

# create confusion matrix
confusionMatrix(pred.test.prune, test.data$hospital_death, positive = "1")
```
```{r testingWithNewCutoff}
pred.prob = predict(cv.ct$finalModel, test.data, type = 'prob')
head(pred.prob)
cutoff <- 0.25 # if proportion of occurrences of class "yes" > cutoff then predicted label = "yes"
pred.test.cutoff <- ifelse(pred.prob[,2] >= cutoff, "1", "0")
confusionMatrix(as.factor(pred.test.cutoff), as.factor(test.data$hospital_death), positive = "1")
```
## Fitting Tree on Non Imputed Instances to compare results

```{r normalData}
rm(list = ls())
source("/Users/jacobmitchell/Documents/STAT6440/functions/myfunctions.R")
hospital <- read.csv("/Users/jacobmitchell/Documents/STAT6440/csvs/dataset.csv")
hospital$hospital_death <- as.factor(hospital$hospital_death)
levels(hospital$hospital_death) <- c("0", "1")
hospital = subset(hospital, select = -c(apache_4a_hospital_death_prob, apache_4a_icu_death_prob))
vals <- unique(hospital$ethnicity)
hospital$ethnicity <- as.factor(hospital$ethnicity)
hospital$ethnicity <- ifelse(hospital$ethnicity == "Caucasian", 1, hospital$ethnicity)
hospital$gender <- as.factor(hospital$gender)
hospital$gender <- ifelse(hospital$ethnicity == "M", 0, hospital$gender)
hospital$gender <- ifelse(hospital$ethnicity == "F", 1, hospital$gender)
hospital$icu_admit_source <- as.factor(hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Accident & Emergency", 0, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Operating Room / Recovery", 1, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Floor", 2, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Other Hospital", 3, hospital$icu_admit_source)
hospital$icu_admit_source <- ifelse(hospital$icu_admit_source == "Other ICU", 4, hospital$icu_admit_source)
hospital$icu_stay_type <- as.factor(hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "admit", 0, hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "readmit", 1, hospital$icu_stay_type)
hospital$icu_stay_type <- ifelse(hospital$icu_stay_type == "transfer", 2, hospital$icu_stay_type)
hospital$icu_type <- as.factor(hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "Med-Surg ICU", 0, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "MICU", 1, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "CTICU", 2, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "CCU-CTICU", 3, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "Neuro ICU", 4, hospital$icu_type)
hospital$icu_type <- ifelse(hospital$icu_type == "SICU", 5, hospital$icu_type)
hospital$apache_3j_bodysystem <- as.factor(hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Cardiovascular", 0, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Metabolic", 1, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Nuerological", 2, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Gastrointestinal", 3, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Metabolic", 4, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Musculoskeletal/Skin", 5, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Respiratory", 6, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Trauma", 7, hospital$apache_3j_bodysystem)
hospital$apache_3j_bodysystem <- ifelse(hospital$apache_3j_bodysystem == "Genitourinary", 8, hospital$apache_3j_bodysystem)
hospital$apache_2_bodysystem <- as.factor(hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Cardiovascular", 0, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Metabolic", 1, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Nuerological", 2, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Gastrointestinal", 3, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Metabolic", 4, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Musculoskeletal/Skin", 5, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Respiratory", 6, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Trauma", 7, hospital$apache_2_bodysystem)
hospital$apache_2_bodysystem <- ifelse(hospital$apache_2_bodysystem == "Genitourinary", 8, hospital$apache_2_bodysystem)
```

```{r looking}
library(tidyr)
hospital = subset(hospital, select = -c(X))
hospital <- drop_na(hospital)
```

```{r more}
RNGkind(sample.kind = "Rounding")
set.seed(0) ## set seed so that you get same partition each time
p2 <- partition.2(hospital, 0.7) ## creating 70:30 partition
training.data <- p2$data.train
test.data <- p2$data.test
```

```{r fitting}
library(rpart)
library(rpart.plot)
library(caret)
set.seed(0)
train_control <- trainControl(method="cv", number=10)
cv.ct <- train(hospital_death ~ . , data = training.data, method = "rpart",trControl = train_control, tuneLength = 10,metric="Kappa")
```

```{r plotsAndSuch}
print(cv.ct)
plot(cv.ct)
```

```{r finalModelSmall}
cv.ct$finalModel
prp(cv.ct$finalModel, type = 1, extra = 1, under = TRUE, split.font = 2, varlen = -10)
```

```{r confusionMatrixSmall}
pred.test.prune = predict(cv.ct$finalModel, test.data, type = 'class')
# create confusion matrix
confusionMatrix(pred.test.prune, test.data$hospital_death, positive = "1")
```

```{r new method}
pred.prob = predict(cv.ct$finalModel, test.data, type = 'prob')
head(pred.prob)
cutoff <- 0.25 # if proportion of occurrences of class "yes" > cutoff then predicted label = "yes"
pred.test.cutoff <- ifelse(pred.prob[,2] >= cutoff, "1", "0")
confusionMatrix(as.factor(pred.test.cutoff), as.factor(test.data$hospital_death), positive = "1")
```
## Discussion of Results
The decision tree with 